### 2017-08-30 status.tex
added some TODOs

\begin{document}
\section{Introduction}
This section describes the current status of the project. In particular, is states the two main models we have and discusses limitations and other problems of both.

@@

Next, we need a model for the block sizes, i.e. the $d_k$s. The longer it takes for a block to arrive, the more transactions and therefore the more transaction data it should accumulate. Let $D(\tau)$ denote the total accumulated transaction data in the block chain. Then $D(\tau)$ we assume that $D(\tau)$ is a nondecreasing L\'{e}vy process, i.e. that $\{D(\tau): \tau > 0\}$ is continuous in $\tau$ with independent, stationary increments. There are many options here, but a convenient L\'{e}vy process is the gamma process, defined by
\begin{align*}
D(\tau + \delta) - D(\tau) &\sim \mathrm{G}(\Psi(\tau + \delta) - \Psi(\tau), \theta)
\end{align*}
for any time $\tau>0$ and increment $\delta > 0$ where $\Psi(\tau) = \int_0^\tau\psi(u)du$, and $\psi(u)>0$ is the intensity function of the  process. We use the shape/rate parameterization of the gamma distribution so that $\mathrm{E}[D(\tau + \delta) - D(\tau)] = [\Psi(\tau + \delta) - \Psi(\tau)]/\theta$. To fully specify this process model we need to specify the intensity function, but for now we will leave that aside. Given the gamma process for $D(\tau)$, we can define the block size process model. Let $D_k = D(\tau_k)$, $d_k = D_k - D_{k-1}$, and $\psi_k = \Psi(\tau_k)$ for $k=1,2,\dots,K$ where we set $D_0 = D(0) = 0$. Then we assume
\begin{align}
d_k | \psi_k, \theta &\stackrel{ind}{\sim}\mathrm{G}(\theta\psi_k, \theta) && \mbox{(block size process model)}\label{eq:size}
\end{align}
for $k=1,2,\dots,K$. We parameterize the intensity function and the gamma process so that $\mathrm{E}[d_k|\psi_k,\theta] = \psi_k$, so that the intensity function defines the mean of $d_k$'s gamma distribution.

Equations \eqref{eq:error} - \eqref{eq:size} define the block size model. To complete this model we must specify the intensity function, $\psi(u)$, and priors for all of the unknown parameters. In the original version of the paper we only considered a constant intensity $\psi(u) = \gamma > 0$, though we can also define an increasing intensity $\psi(u) = e^{\beta_0 + u\beta_1}$. In principle we can define any intensity function of the form $\psi(u) = e^{\bm{z}(u)'\bm{\beta}}$ where $\bm{z}(u)$ is a vector of basis functions of time, and $\bm{\beta}$ is a vector of regression coefficients. However, we need to be able to compute the integral $\int_{\tau}^{\tau + \delta}\psi(u)du$ in closed form, which restricts the form of $\psi(u)$. One possibility that I have not explored that may be more flexible is $\psi(u) = \sum_{p=1}^P\alpha_pu^{\beta_p - 1}$ with $\alpha_p, \beta_p>0$ for $p=1,2,\dots,P$.

@@

As an alternative, what we could do is stick with the model we have above and fit a separate model using block sizes (in MB) and block transaction counts in order to estimate the size of a typical transaction. Then we can compare the distribution of transaction sizes to the difference between the oberved block sizes and the cap in order to predict when the cap has been reached. We can do the same thing with the {\it predicted} block sizes, which we can obtain from the posterior predictive distribution of the block size model. This is substantially easier to do.
\end{document}

>>>>

\begin{document}
\section{Introduction}
This note describes the current status of the project. In particular, is states the first model we used and discusses its limitations. Then it discusses what to do about those limitations, including whether we should use the second model.

@@

Next, we need a model for the block sizes, i.e. the $d_k$s. The longer it takes for a block to arrive, the more transactions and therefore the more transaction data it should accumulate. Let $D(\tau)$ denote the total accumulated transaction data in the block chain. Then $D(\tau)$ we assume that $D(\tau)$ is a nondecreasing L\'{e}vy process, i.e. that $\{D(\tau): \tau > 0\}$ is continuous in $\tau$ with independent, stationary increments. There are many options here, but a convenient L\'{e}vy process is the gamma process, defined by
\begin{align*}
D(\tau + \delta) - D(\tau) &\sim \mathrm{G}(\Psi(\tau + \delta) - \Psi(\tau), \theta)
\end{align*}
for any time $\tau>0$ and increment $\delta > 0$ where $\Psi(\tau) = \int_0^\tau\psi(u)du$, and $\psi(u)>0$ is the intensity function of the  process. We use the shape/rate parameterization of the gamma distribution so that $\mathrm{E}[D(\tau + \delta) - D(\tau)] = [\Psi(\tau + \delta) - \Psi(\tau)]/\theta$. To fully specify this process model we need to specify the intensity function, but for now we will leave that aside. Given the gamma process for $D(\tau)$, we can define the block size process model. Let $D_k = D(\tau_k)$, $d_k = D_k - D_{k-1}$, and $\psi_k = \Psi(\tau_k)$ for $k=1,2,\dots,K$ where we set $D_0 = D(0) = 0$. Then we assume
\begin{align}
d_k | \psi_k, \theta &\stackrel{ind}{\sim}\mathrm{G}(\theta\psi_k, \theta) && \mbox{(block size process model)}\label{eq:size}
\end{align}
for $k=1,2,\dots,K$. We parameterize the intensity function and the gamma process so that $\mathrm{E}[d_k|\psi_k,\theta] = \psi_k$, so that the intensity function is $\theta\psi(u)$ and $\psi(u)$ defines the mean of $d_k$'s gamma distribution.

Equations \eqref{eq:error} - \eqref{eq:size} define the block size model. To complete this model we must specify the intensity function, $\psi(u)$, and priors for all of the unknown parameters. In the original version of the paper we only considered a constant intensity $\psi(u) = \gamma > 0$, though we can also define an increasing intensity $\psi(u) = e^{\beta_0 + u\beta_1}$. In principle we can define any intensity function of the form $\psi(u) = e^{\bm{z}(u)'\bm{\beta}}$ where $\bm{z}(u)$ is a vector of basis functions of time, and $\bm{\beta}$ is a vector of regression coefficients. However, we need to be able to compute the integral $\int_{\tau}^{\tau + \delta}\psi(u)du$ in closed form, which restricts the form of $\psi(u)$. One possibility that I have not explored that may be more flexible is $\psi(u) = \sum_{p=1}^P\alpha_pu^{\beta_p - 1}$ with $\alpha_p, \beta_p\geq 0$ for $p=1,2,\dots,P$. In this case, then $\Psi(\tau) = \int_0^\tau\psi(u)du = \alpha_p\tau^{\beta_p}/\beta_p$ where $\tau^0/0$ is defined to be $\log(\tau)$. If $\beta_p < 1$ then the $p$th term of $\psi(u)$ decays in $u$, while if $\beta_p > 1$ it increases in $u$ and if $\beta_p = 1$ it is constant in $u$.

@@

As an alternative, what we could do is stick with the model we have above and fit a separate model using block sizes (in MB) and block transaction counts in order to estimate the size of a typical transaction. Then we can compare the distribution of transaction sizes to the difference between the oberved block sizes and the cap in order to predict when the cap has been reached. We can do the same thing with the {\it predicted} block sizes, which we can obtain from the posterior predictive distribution of the block size model. This is substantially easier to do.

The transaction data model might look like this. Suppose $x_i$ denotes the amount of data for transaction $i$, then we assume $x_i \stackrel{iid}{\sim}\mathrm{G}(\alpha, \beta)$. But we do not observed the $x_i$s. Instead we observe the $d_k$, the amount of data in block $k$, and $n_k$, the number of transactions in block $k$. The model on the $x_i$s implies then that $d_k \stackrel{ind}{\sim} \mathrm{G}(\alpha n_k, \beta)$. So we can fit this model and get parameter estimates for $\alpha$ and $\beta$, and in particular compute $P(x_i > x)$ --- the probability that a given transaction is size $x$ or large for any $x$. When $x$ is the difference between the block size cap and an observed block size, then this is an estimate of the probability that there was block overflow.

\section{What to do next}
In short, this is what I want to do in order to get this paper into a form that we can submit:
\begin{enumerate}
\item Get updated data from the block chain. (Can we easily do this?)
\item Fit transaction size models. The model I mentioned in the last paragraph of the previous section assumes that transactions {\it sizes} have been constant, on average, over time, and any differences in block sizes is due to the number of transactions. We can investigate this pretty easily.
\item Fit several versions of the block size model --- basically with different choices of $\psi(u)$. 
\item Compare predictions of the block size model to reality --- do some one step ahead predictions to see how well we predict block sizes. We can do this in different regions of the blockchain --- e.g. with the old data we have, and also with up-to-the-minute data. 
\item Use 1) and 2) in order to estimate the probability of transaction spillover in previous blocks, and new blocks. 
\item Write everything up.
\end{enumerate}

Getting these models implemented and fit shouldn't be much work for me. Everything will be in Stan, so it'll mostly be a matter of pressing buttons. The most time consuming part will be writing everything up, including getting references to the relevant literature.
\end{document}
